---
title: "HW5"
author: "Nihar Harkare"
date: "2024-02-26"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mosaic)
library(tidyverse)
```

# Problem 1

## Null Hypothesis:

The baseline probability that any trade will be flagged by the SEC's algorithm is 2.4%

## Test Statistic:

70/2021 trades being flagged by the SEC's algorithm, or `r round((70/2021), 4)`.

## Plot of the probability distribution:

```{r echo=FALSE}
sim_suspicious_trades = do(100000)*nflip(n=2021, prob=0.024)

ggplot(sim_suspicious_trades) + 
  geom_histogram(aes(x=nflip), binwidth=0.5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Suspicious Trades", x = "Number of Suspicious Trades", y = "Frequency")

p_suspicious_trades = sum(sim_suspicious_trades >= 70)/100000
```

## P-Value:

`r p_suspicious_trades`

## Conclusion:

Assuming the null hypothesis is true, there is a `r (p_suspicious_trades * 100)` percent chance that we would get a test statistic of `r (70/2021)` or higher. Since p (`r p_suspicious_trades`) \< 0.05, there is a statistically significant chance that SEC's algorithm is also flagging illegal market activity.

# Problem 2

## Null Hypothesis:

The baseline probability that any restaurant inspection will result in a health code violation by the Health Department is 3%

## Test Statistic:

8/50 inspections resulting in a health code violation by the Health Department, or `r (8/50)`.

## Plot of the probability distribution:

```{r echo=FALSE}
sim_healthcode_violations = do(100000)*nflip(n=50, prob=0.03)

ggplot(sim_healthcode_violations) + 
  geom_histogram(aes(x=nflip), binwidth=0.5, fill = "orange", color = "black") +
  labs(title = "Distribution of Health Code Violations", x = "Number of Violations", y = "Frequency")

p_healthcode_violations = sum(sim_healthcode_violations >= 8)/100000
```

## P-Value:

`r p_healthcode_violations`

## Conclusion:

Assuming the null hypothesis is true, there is a `r (p_healthcode_violations * 100)` percent chance that we would get a test statistic of `r (8/50)` or higher. Since p (`r p_healthcode_violations`) \< 0.05, there is a statistically significant chance that Gourmet Bites' rate of health code violations is significantly higher than the citywide average of 3%.

# Problem 3

## Part A

```{r echo=FALSE}
letter_frequencies = read_csv("letter_frequencies.csv", show_col_types = FALSE)
brown_sentences = readLines("brown_sentences.txt")

brown_sentences = paste(brown_sentences, collapse = " ")
clean_sentences = gsub("[^A-Za-z.]", "", brown_sentences)
clean_sentences = toupper(clean_sentences)

letter_counts = table(strsplit(clean_sentences, NULL))
letter_counts_df = as.data.frame(table(strsplit(clean_sentences, NULL)))
colnames(letter_counts_df) = c("Letter", "Count")

letter_counts_no_per = subset(letter_counts_df, Letter != ".")

letter_counts_sentences = subset(letter_counts_df, Letter != ".")
letter_counts_sentences$Count = letter_counts_sentences$Count / letter_counts_df$Count[1]
letter_counts_sentences$FrequencyPerSentence = letter_counts_sentences$Count / sum(letter_counts_sentences$Count)

total_letters = sum(letter_counts_no_per$Count)
expected_letter_counts = merge(letter_frequencies, letter_counts_no_per, by = "Letter")
expected_letter_counts$Expected_Count = expected_letter_counts$Probability * total_letters
expected_letter_counts$Expected_Count_Sentence = expected_letter_counts$Expected_Count / letter_counts_df$Count[1]

observed_counts = letter_counts_sentences$Count
expected_counts = expected_letter_counts$Expected_Count_Sentence
chi_square_stat = sum((observed_counts - expected_counts)^2 / expected_counts)

boot_actual_counts <- replicate(10000, letter_counts_no_per$Count[sample(seq_along(letter_counts_no_per$Count), replace = TRUE)])

chi_square_stats <- numeric(10000)

for (i in seq_along(chi_square_stats)) {
  boot_actual_counts_i <- boot_actual_counts[, i]
  chi_square_stats[i] <- sum((letter_counts_no_per$Count - boot_actual_counts_i)^2 / letter_counts_no_per$Count)
}
```

`r chi_square_stat`

## Part B

```{r echo=FALSE}
sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

calculate_chi_squared = function(sentence, freq_table) {
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  
  clean_sentence = gsub("[^A-Za-z]", "", sentence)
  clean_sentence = toupper(clean_sentence)
  
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)
}

p_sentence = numeric(10)

for (i in seq_along(sentences)) {
  chi_sentence = calculate_chi_squared(sentences[i], letter_frequencies)
  p_sentence[i] = sum(chi_square_stats >= chi_sentence) / 10000
}

#I appear to have done something horribly wrong
p_vals_df <- data.frame(Sentence = sentences, P_Value = round(p_sentence, 3))
print(p_vals_df)

min_p_value_row <- p_vals_df[which.min(p_vals_df$P_Value), ]
min_sentence <- min_p_value_row$Sentence
min_p_value <- min_p_value_row$P_Value

cat(paste("According to the table, the sentence written by an LLM is:", min_sentence, "with a p-value of", min_p_value))
```